---
title: "Lab and Homework 07 - Fitting models`"
author: "Louis Quigley"
date: "03/30/23"
output: pdf_document
---

### Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse)
library(tidymodels)
```

## Exercises

### Exercise 1

```{r import-data}
fluxes <- read_csv("data/us-mms-simple.csv")
```

```{r fix-dates, eval = T}
fluxes <- fluxes %>% 
  mutate(date = mdy_hm(date))
head(fluxes)
```

### Exercise 2

```{r subset-fluxes, eval = T}
fluxes_subset <- fluxes %>% 
  filter(date > "2020-07-01" & date < "2020-07-08")
head(fluxes_subset, 10) 
```

### Exercise 3

```{r CO2_exploratory_viz}
ggplot(fluxes_subset, aes(incoming_radiation, fco2)) +
  geom_point() +
  labs(
    title = "CO2 Flux vs. Incoming Radiation",
    x = "Incoming Radiation",
    y = "CO2 Flux"
  ) +
  theme_minimal()
```


### Exercise 4

```{r add_line}
ggplot(fluxes_subset, aes(incoming_radiation, fco2)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  labs(
    title = "CO2 Flux vs. Incoming Radiation",
    x = "Incoming Radiation",
    y = "CO2 Flux"
  ) +
  theme_minimal()
```
The grey band shows the uncertainty of the fitted line.

### Exercise 5

```{r fit-a-model, eval = F}
linear_reg() %>% 
  set_engine("lm") %>% 
  fit(fco2 ~ incoming_radiation, data = fluxes_subset) %>% 
  tidy()
```
1. The "intercept" estimate is the value of CO2 flux when the incoming radiation equals 0.  
2. The slope is the estimated change in CO2 flux as incoming solar radiation increases.  
std.error is the uncertainty of estimated coefficients (and should be tested at a local healthcare facility)

### Exercise 6

```{r create-season, eval = T}
fluxes <- fluxes %>% 
    filter(date > "2020-01-01" & date < "2021-01-01") %>%  ## THIS LINE IS NEW 
     mutate(season = ifelse(date < "2020-04-01", "winter", NA),
            season = ifelse(date >= "2020-04-01" & date < "2020-07-01", "spring", season),
            season = ifelse(date >= "2020-07-01" & date < "2020-10-01", "summer", season),
            season = ifelse(date >= "2020-10-01" & date <= "2020-12-31", "fall", season)) %>% 
     mutate(winter = ifelse(season == "winter", 1, 0))

# ifelse syntax is: ifelse( logical operation, value if TRUE, value if FALSE )

```


### Exercise 7

```{r fit-a-model-2, eval = T}
linear_reg() %>% 
  set_engine("lm") %>% 
  fit(fco2 ~ factor(winter), data = fluxes) %>% 
  tidy()
```
The estimate column shows the difference in CO2 flux when it is winter versus not winter.  
The slope tells us by how much the CO2 flux is changing between winter and not winter.  
There are two possible values for fCO2 because it's calculating either winter or not winter values. Exercise 6 gives us multiple dates for fCO2 values. 

### Exercise 8

```{r fit-a-model-3, eval = F}
linear_reg() %>% 
  set_engine("lm") %>% 
  fit(fco2 ~ factor(season), data = fluxes) %>% 
  tidy()
```
the estimate variable is exaplaining the estimated value of fCO2 by different seasons starting with the fall as the y-intercept.  
There are four different possible values because this table models four seasons.  
Fall is the baseline or y-intercept because it is listed first and r defaults to the first datapoint listed as the baseline.  
### Exercise 9

Remove this text, and add your answer for Exercise 8 here.

### Exercise 10

Remove this text, and add your answer for Exercise 8 here.

### Exercise 11

Remove this text, and add your answer for Exercise 8 here.

### Exercise 12

Remove this text, and add your answer for Exercise 8 here.
